Binning Strategies and Related Loss for Large Data

### 2015-03-29 Paper.Rnw
Add potential freq loss to section 4.2

\ktm{ The net spatial loss is perhaps more interpretable if it is viewed in comparison with the spatial loss from a ``worst-case scenario''. The largest displacement of points will occur if we place all points into a single bin.  While the total spatial loss for single bin could be made arbitrarily large if the bin center is placed far away from the data values, we will consider a reasonable specification of a bin with boundaries that span only the range of the data in each dimension; thus centered at  $( (x_{(1)}-x_{(n)})/2$ , $(y_{(1)}-y_{(n)})/2 )$.  The total spatial loss for this ``worst-case'' bin will be referred to as the \textit{potential spatial loss}, $L^S_p$, which is defined as }
\begin{eqnarray}\label{potspatloss}
L^S_p = \sum_{i=1}^{n} \sqrt{ \left(x_i-(x_{(1)}-x_{(n)})/2\right)^2 + \left(y_i-(y_{(1)}-y_{(n)})/2\right)^2 }
\end{eqnarray} 
%
\ktm{ Thus the net spatial loss as a proportion of potential spatial loss, $L^S_{net}/L^S_p$, will be used as the primary metric for assessing spatial loss incurred from binning.}

@@

>>>>

\ktm{ The net spatial loss is a euclidean distance, but requires a reference scale on which the magnitude may be interpretable. We will use the spatial loss from a ``worst-case'' binning scenario as a reference value. The largest displacement of points will occur if we place all points into a single bin.  While the total spatial loss for single bin could be made arbitrarily large if the bin center is placed far away from the data values, we will consider a reasonable specification of a bin with boundaries that span only the range of the data in each dimension; thus centered at  $( (x_{(1)}-x_{(n)})/2$ , $(y_{(1)}-y_{(n)})/2 )$.  The total spatial loss for this ``worst-case'' bin will be referred to as the \textit{potential spatial loss}, $L^S_p$, which is defined as }
\begin{eqnarray}\label{potspatloss}
L^S_p = \sum_{i=1}^{n} \sqrt{ \left(x_i-(x_{(1)}-x_{(n)})/2\right)^2 + \left(y_i-(y_{(1)}-y_{(n)})/2\right)^2 }
\end{eqnarray} 
%
\ktm{ Thus the net spatial loss, as a proportion of potential spatial loss, $L^S_{net}/L^S_p$, will be used as the primary metric for assessing spatial loss incurred from binning.}

@@

\ktm{ As with spatial binning, there will be a ``worst-case'' scenario for frequency binning that will provide context to the magnitude of the total frequency loss.  If we elect to use a single frequency bin then all tiles of the binned scatterplot will be rendered with the same shade; thus the shade would represent the average bin density, $\bar{c} =  (\sum_{\ell=1}^{L} c_\ell) /L$.  We define the total frequency loss in the single shade scenario as the \textit{protential frequency loss}, $L^F_p$,  where
%
\begin{eqnarray}\label{potfreqloss}
L^F_p = \sum_{\ell=1}^{L} (c_\ell-\bar{c})^2.
\end{eqnarray}  
%
Thus total frequency loss after binning will be interpreted through the proportion of potential frequency loss, $L^F / L^F_p$.  Similarly we define the \textit{protential log frequency loss}, $L^{\log F}_p$, as 
%
\begin{eqnarray}\label{potlogfreqloss}
L^{\log F}_p = \sum_{\ell=1}^{L} (\log(c_\ell+1)-\log(\bar{c}+1))^2,
\end{eqnarray}  
% 
which can then be used to find the proportion of potential log frequency loss, $L^{\log F} / L^{\log F}_p$, in cases where log frequency binning was conducted. } %end ktm{} red
